\documentclass[10pt,english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{geometry}
\geometry{verbose,tmargin=1.5in,bmargin=1.5in,lmargin=1.5in,rmargin=1.5in}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}

\makeatletter
\usepackage{enumitem}
\newlength{\lyxlabelwidth}

\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}

%\usepackage{txfonts}

\usepackage{microtype}

\usepackage{calc}
\usepackage{enumitem}
\setenumerate{leftmargin=!,labelindent=0pt,itemindent=0em,labelwidth=\widthof{\ref{last-item}}}

\makeatother

\usepackage{babel}
\begin{document}
\noindent \begin{center}
\textbf{\large{}MATH 146 - Assignment 1}\\
\textbf{\large{}Chris Ji 20725415}
\par\end{center}{\large \par}
\medskip{}

\begin{enumerate}
\item \begin{enumerate}
    \item The distance of $C$ can be found easily by comparing each code word. \\\\
    \begin{tabular}{|c|c|c|c|c|}
\hline
Code word & $c_1$&$c_2$&$c_3$&$c_4$ \\ \hline
$c_1$      & - & 3 & 3 & 4  \\
$c_2$      & 3 & - & 4 & 3    \\      
$c_3$      & 3 & 4 & - & 3    \\ 
$c_4$      & 4 & 3 & 3 & -    \\
\hline
\end{tabular}\\ \\
Then clearly the distance of $C$ is 3.

    \item \begin{align*}
        P(r|c_1)=0.1^10.9^4= 0.06561\\
        P(r|c_2)=0.1^40.9^1= 0.00009\\
        P(r|c_3)=0.1^20.9^3= 0.00729\\
        P(r|c_4)=0.1^30.9^2= 0.00081
        \end{align*}
        so IMLD would decode this to $c_1$. 
    
    
    \item \begin{align*}
        P(c_1|r)=P(r|c_1)\frac{P(c_1)}{P(r)}=0.1^10.9^4\frac{0.1}{P(r)}=\frac{0.006561}{P(r)}\\
        P(c_2|r)=P(r|c_1)\frac{P(c_1)}{P(r)}=0.1^40.9^1\frac{0.2}{P(r)}=\frac{0.000018}{P(r)}\\
        P(c_3|r)=P(r|c_1)\frac{P(c_1)}{P(r)}=0.1^20.9^3\frac{0.2}{P(r)}=\frac{0.001458}{P(r)}\\
        P(c_4|r)=P(r|c_1)\frac{P(c_1)}{P(r)}=0.1^30.9^2\frac{0.5}{P(r)}=\frac{0.000405}{P(r)}
        \end{align*}
        so MED would decode this to $c_1$. 
    
    \item \begin{align*}
        P(r|c_1)=0.4^10.6^4= 0.05184\\
        P(r|c_2)=0.4^40.6^1= 0.01536\\
        P(r|c_3)=0.4^20.6^3= 0.03456\\
        P(r|c_4)=0.4^30.6^2= 0.02304
        \end{align*}
        so IMLD would decode this to $c_1$. 


    \item \begin{align*}
        P(c_1|r)=P(r|c_1)\frac{P(c_1)}{P(r)}=0.4^10.6^4\frac{0.1}{P(r)}=\frac{0.005184}{P(r)}\\
        P(c_2|r)=P(r|c_1)\frac{P(c_1)}{P(r)}=0.4^40.6^1\frac{0.2}{P(r)}=\frac{0.003072}{P(r)}\\
        P(c_3|r)=P(r|c_1)\frac{P(c_1)}{P(r)}=0.4^20.6^3\frac{0.2}{P(r)}=\frac{0.006912}{P(r)}\\
        P(c_4|r)=P(r|c_1)\frac{P(c_1)}{P(r)}=0.4^30.6^2\frac{0.5}{P(r)}=\frac{0.01152}{P(r)}
        \end{align*}
        so MED would decode this to $c_4$ 
    
\end{enumerate}

\pagebreak
\item If we set $p'=1-p$, then 
    \begin{alignat*}{4}
        \frac{q-1}{q}&\leq p&&\leq1\\
        \Rightarrow 1-1&\leq p' &&\leq1-\frac{q-1}{q}\\
        \Rightarrow 0&\leq p'&&\leq\frac{1}{q}
    \end{alignat*}
Then clearly $\frac{1}{q}\leq\frac{q-1}{q}$ is true for all $q<0$ and $q\geq2$, and $q$ has to be at least 2, or else our alphabet has $1$ or less symbols, in which case it is always perfect (1 symbol), or doesn't exist (less than 1 symbol). 
    


\pagebreak
\item Suppose $c\in C$ was sent, and $r$ was received. Consider if at most $t$ errors occurred. Then $d(r,c)\leq t$. If $c'$ is any other codeword, then $d(r,c')\geq d(c',c)-d(c,r)\geq 2t+s+1-t=t+s+1>t$, and so $c$ is the minimum distance codeword from $r$, and the decoder correctly corrects $r$ to $c$. \\ 
Now consider if $e$ errors were made. If $e=0$, then $r$ is accepted. If $1\leq e\leq s+t$, then $1\leq d(r,c)\leq s+t$, and so $r\not\in C$ (since the minimum distance between any two words in $C$ is $2t+s+1$), and is rejected. Note that there exist two words $c_1,c_2\in C$ such that $d(c_1,c_2)=2t+s+1$. Then if $c_1$ is sent, and $c_2$ is received, we have $e=2t+s+1>t+s$, and $c_2$ is incorrectly accepted.  


\pagebreak
\item \begin{enumerate}
    \item Let $x=(x_1,\ldots,x_n)$, and let $y=(y_1,\ldots,y_n), z=(z_1,\ldots,z_n)$ be any two words in our code, $C$. Consider if $y$ and $z$ were different at some index $i$. Then if $x_i=0$, $y_i+0\not\equiv z_i+0\quad\text{mod }2$ still holds true. If $x_i=1$, then $y_i+1\not\equiv z_i+1\quad\text{mod }2$ is also still true. Consider if $y$ and $z$ were the same at some index $j$. Then, if $x_j=0$, $y_j+0\equiv z_j+0\quad\text{mod }2$ is still true. If $x_j=1$, then $y_j+1\equiv z_j+1\quad\text{mod }2$ is also still true. Hence,y $d(y,z)=d(y+x,z+x)$. Since this is true for any two words in our code, clearly $d(C)=d(C+x)$. 
    
    
    \item $C=\{111110000,101010111,010101011,000000000\}$ is a $[9,4]-$code with distance 5.

    \item Let $x,y,z$ be the three words in our code. Consider an $[n,3]-code$ such that $d>\frac{2n}{3}$. Clearly $d(x,y)\leq n$ (without loss of generality by choosing $x$ and $y$), as the words can't differ in more symbols than are in it. Then 
    \begin{align}
    n\geq d(x,y)&>\frac{2n}{3}\\ 
    n\geq d(y,z)&>\frac{2n}{3}\\ 
    n\geq d(x,z)&>\frac{2n}{3} 
    \end{align}
    Subtracting $(2)$ from $(1)$, we get $0\geq d(x,y)-d(y,z)>0$, or $0>0$, which is a clearly a contradiction. Thus $d$ has to be less than or equal to $\frac{2n}{3}$.
    
    
    
\end{enumerate}
\end{enumerate}

\end{document}
